Memory Safety: A Systematic Literature Review

### 2024-04-01 paper.tex
Change paperpile bib to export based on tag not folder
Working on SLR protocal section

\documentclass[sigconf]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}

@@

In the ever-evolving landscape of software development, the reliability and security of computer
systems stand as paramount concerns for all parties involved. Modern software is constructed by
building ever more complex abstractions. Each abstraction forms a layer that the next one it built
upon. Thus, if we aim to have a secure system we must peel back all the layers and tackle one of the
fundamental abstractions in computer science the programming language. Programming languages that
provided and enforce memory safety eliminate whole classes of bugs such as buffer overflows,
dangling pointers, and memory leaks which have been implicated in a myriad of security
vulnerabilities and system crashes.

This Systematic Literature Review (SLR) aims to provide a comprehensive and up-to-date analysis of
the existing body of research surrounding memory safety. By systematically reviewing and
synthesizing the wealth of knowledge accumulated in this field, we seek to uncover the current state
of understanding, identify emerging trends, and highlight gaps in our knowledge that warrant further
investigation.

@@

Before we start researching we must first confirm the need for a SLR. The recent report released by
ONCD~\cite{United_States_Gov2024-pp} has conveniently done this job for us by compiling a report
detailing the need for research in the domain of memory safety. While the ONCD report detailed a two
pronged approach for this SLR we will be focusing on a memory safe programming language,
specifically Rust.

@@

\begin{itemize}
    \item \textbf{RQ1:} How can \lstinline{unsafe} Rust derived from C be converted to safe Rust?
    \item \textbf{RQ2:} How can we build Rust bindings into core abstractions?
    \item \textbf{RQ3:} How can we better find vulnerabilities in \lstinline{unsafe} blocks
\end{itemize}

@@

@@

@@

@@

@@

@@

@@

In this section we address the proposed research questions.

@@

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{paperpile.bib}

>>>>

\documentclass[sigconf]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}
\pagestyle{plain} %to show page numbers

@@

In the ever-evolving landscape of software development, the reliability and security of computer
systems stands as a paramount concern for all parties involved. Modern software is constructed by
building ever more complex abstractions, one on top of the other. Thus, if we aim to have a secure
system we must start to peel back all the layers and tackle one of the fundamental abstractions in
computer science, the programming language. Programming languages that provided and enforce memory
safety eliminate whole classes of bugs such as buffer overflows, dangling pointers, and memory leaks
which have been implicated in a myriad of security vulnerabilities and system crashes.

The migration of legacy C codebases to memory-safe programming languages like Rust is of paramount
importance in modern software development, aiming to mitigate common security vulnerabilities and
memory-related errors. This paper presents a systematic literature review (SLR) focusing on
strategies and methodologies for converting C code to Rust while minimizing the utilization of
unsafe blocks. Our study aims to provide a comprehensive overview of existing research, identify
gaps, and suggest future directions in this domain. Through a rigorous search process, we
synthesized relevant studies and extracted key findings to offer insights into effective approaches
for ensuring memory safety during code migration.

@@

Before we start researching we must first confirm the need for a SLR. The recent report released by
ONCD~\cite{United_States_Gov2024-pp} has conveniently done this job for us by compiling a report
detailing the need for research in the domain of memory safety. While the ONCD report detailed a two
pronged approach for this SLR we will be focusing on a memory safe programming language,
specifically Rust. While there are many memory safe programming languages available for software
developers to use Rust is the only language that has support in kernel
space~\cite{The_kernel_development_community_undated-iw} thus is a primary candidate to replace the
aging C programming language.

@@

\begin{itemize}

    \item \textbf{RQ1:} What are the existing methodologies and strategies for converting C code to
      Rust?
    \item \textbf{RQ2:} How do these methodologies address memory safety concerns and minimize the
      use of \lstinline{unsafe} blocks?
    \item \textbf{RQ3:} What are the challenges and limitations encountered in the conversion process?
\end{itemize}

@@

We employed a robust multi-step search strategy across three digital databases, ACM Digital Library,
IEEE Xplore, and Google Scholar in order to find all the current research that has been done. These
databases cover a wide range of current software engineering and computer science literature. We
leveraged the advanced search features of both the ACM Digital Library and IEEE Xplore to search the
titles and abstracts for keywords using boolean search operators. We then searched Google Scholar
using the same keywords and operators sorted by relevance and took the top 250 results.

@@

The search criteria for each database is shown in table~\ref{tab:keywords}.  First, we searched
titles and abstracts in the ACM Digital Library and IEEE Xplore with the keywords ``C rust convert
unsafe'' from May 2015 to the present~\footnote{This was the release date of rust 1.0}. Then we
searched on google Google Scholar using the same keywords and dates in order to to cast as wide a
net as possible.

\begin{table}
\begin{tabular}{|| p{3cm}| p{5cm} ||}
 \hline
 Database & Query \\
 \hline\hline
 ACM Digital Library & Title:(C conver* unsafe rust) AND Abstract:(C conver* unsafe rust)  \\
 IEEE Xplore & TBD \\
 Google Scholar & TBD \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

Search keywords: "C to Rust conversion", "memory-safe Rust", "legacy code migration", "unsafe blocks
mitigation", "safe programming practices".

@@

Inclusion criteria: Papers published between 2012 and 2024, written in English, focusing on
converting C code to Rust with a focus on memory safety.

Exclusion criteria: Duplicate studies, non-peer-reviewed articles, unrelated topics.

Manual search: Relevant conference proceedings, journals, and citations of selected papers.

@@

Initial screening based on titles and abstracts.  Full-text assessment for eligibility based on
inclusion and exclusion criteria.  Selection of final studies for data extraction and analysis.

@@

Extraction of relevant data including methodologies, tools, challenges, and outcomes.

@@

Synthesis of key findings and categorization based on common themes.
Identification of trends, gaps, and areas for further investigation.

@@

Presentation of synthesized findings on existing methodologies and strategies.  Discussion on the
effectiveness of approaches in ensuring memory safety and minimizing unsafe blocks.  Analysis of
challenges and limitations encountered in the conversion process.  Comparison of different tools and
techniques used in C to Rust migration.

Implications of the findings for software developers, researchers, and practitioners.
Recommendations for improving existing methodologies and addressing identified gaps.
Suggestions for future research directions in the field of C to Rust conversion.

@@

Summary of key findings and insights obtained from the systematic literature review.
Concluding remarks on the significance of memory-safe programming languages in mitigating security vulnerabilities.
Closing thoughts on the importance of minimizing unsafe blocks during code migration.

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{slr-paper-memory-safety-rust.bib}
